---
title: "Structured Independent Edge Model"
author: "Eric Bridgeford"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIEM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, warning=FALSE, echo=FALSE}
require(graphstats)
require(ggplot2)
require(latex2exp)
```

For a walkthrough including the model, see [model](http://docs.neurodata.io/graphstats/siem/siem.html).

# Example


In our below example, we simulate a graph where p~c1~ = 0.3 and p~c2~ = 0.7:

```{r, fig.width=5, fig.height=3.5}
nv <- 20  # number of vertices in simulated graph
X <- array(0, dim=c(nv, nv))  # simulated graph initialized
com <- array(0, dim=c(nv, nv))  # probability at each edge
Es <- list()
split <- sample(nv^2, nv^2, replace=FALSE)  # sample edges in random order
pedge <- data.frame(community=c(), result=c(), p=c())

Es[[1]] <- split[1:(nv^2/2)]  # first half in edge-community 1
Es[[2]] <- split[(nv^2/2+1):(nv^2)]  # second half in edge-community 2

p <- c(.3, .7)  # probabilities between community 1 and 2 are vastly different

for (i in 1:length(Es)) {
  X[Es[[i]]] <- rbinom(n=length(Es[[i]]), prob=p[i], size=1)
  com[Es[[i]]] <- i
  pedge <- rbind(pedge, data.frame(community=c(i), result=c("true"), p=c(p[i])))
  pedge <- rbind(pedge, data.frame(community=c(i), result=c("sampled"), p=c(sum(X[Es[[i]]])/length(Es[[i]]))))
}

pedge$community <- factor(pedge$community)
ggplot(pedge, aes(x=community, y=p, fill=result)) +
  geom_bar(stat="identity", width=.5, position = "dodge") +
  ggtitle("Comparing True vs. Sim Edge Community, Sample 1")
gs.plot.plot_matrix(com, legend.name = "community", xlabel = "vertex",
                     ylabel = "vertex", title = "Community Each Edge is from, Sample 1", ffactor=TRUE)
gs.plot.plot_matrix(X, legend.name = "connection", xlabel = "vertex",
                     ylabel = "vertex", title = "Graph Simulated from SIEM, Sample 1")
```

## One Sample Test

Given a graph from a single sample, we might be interested in whether two estimators within that single sample differ. For example, we can detect the probability of an edge in one community singificantly exceeding the probability of an edge in another community for our example graph:

```{r, fig.width=5, fig.height=3.5}
model <- gs.siem.fit(X, Es)  # fit siem model
os_pval <- array(0, dim=c(2, 2))
for (i in 1:length(Es)) {
  for (j in 1:length(Es)) {
    os_pval[i, j] <- gs.siem.sample.test(model$pr[i], model$pr[j], model$var[i],
                                         model$var[j], 1, alt='greater')$p
  }
}

gs.plot.plot_matrix(os_pval, legend.name = TeX("$p$-value"), xlabel = TeX("$c_i$"),
                     ylabel = TeX("$c_j$"), title = TeX("$p$-value that $p_{c_i} > p_{c_j}$, Sample 1"),
                     limits=c(0, 1), vfactor=TRUE)
```

as we can see, with a = 0.05, we reject the null in favor of the alternative for the combination p~c2~ > p~c1~, which intuitively makes sense under the parameters we estimated previously as p~c2~ = 0.7 and p~c2~ = 0.3.

## Two Sample Test

In our below example, we simulate a second graph where p~c1~ = 0.5 and p~c2~ = 0.6:

Given a second graph from a different sample, we may be interested in whether two estimators across the 2 samples differ. For example, we might be interested in whether the difference in the estimator of the probability of an edge in c~2~ exceeds the probability of an edge in c~1~ from sample 1 to sample 2. That is, whether d~s1~ = x~c1,1~ - x~c2,1~ > d~s2~ = x~c1,2~ - x~c2,2~:

```{r, fig.height=3.5, fig.width=5}
X2 <- array(0, dim=c(nv, nv))  # simulated graph initialized
com2 <- array(0, dim=c(nv, nv))  # probability at each edge
pedge <- data.frame(community=c(), result=c(), p=c())

p2 <- c(.5, .6)  # probabilities between community 1 and 2 are vastly different

for (i in 1:length(Es)) {
  X2[Es[[i]]] <- rbinom(n=length(Es[[i]]), prob=p2[i], size=1)
  com[Es[[i]]] <- i
  pedge <- rbind(pedge, data.frame(community=c(i), result=c("true"), p=c(p2[i])))
  pedge <- rbind(pedge, data.frame(community=c(i), result=c("sampled"), p=c(sum(X[Es[[i]]])/length(Es[[i]]))))
}

pedge$community <- factor(pedge$community)
ggplot(pedge, aes(x=community, y=p, fill=result)) +
  geom_bar(stat="identity", width=.5, position = "dodge") +
  ggtitle("Comparing True vs. Sim Edge Community, Sample 2")
gs.plot.plot_matrix(com, legend.name = "community", xlabel = "vertex",
                     ylabel = "vertex", title = "Community Each Edge is from, Sample 2", ffactor=TRUE)
gs.plot.plot_matrix(X, legend.name = "connection", xlabel = "vertex",
                     ylabel = "vertex", title = "Graph Simulated from SIEM, Sample 2")
```

```{r}
model2 <- gs.siem.fit(X2, Es)  # fit siem model
gs.siem.sample.test(model$dpr[2, 1], model2$dpr[2, 1], model2$dvar[2, 1],
                    model2$dvar[2, 1], 2, alt='greater')$p
```

and we can see that at a=0.05 we detect a significant difference between the two.
