---
title: "Vertex Nomination via Seeded Graph Matching"
author: "Kemeng Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vn}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, warning=FALSE}
# Dependencies.
suppressMessages(require(graphstats))
suppressMessages(require(igraph))
suppressMessages(require(ggplot2))
```


# Background

Consider two networks on overlapping, non-identical vertex sets.  Given vertices of interest in the first network, we seek to identify the corresponding vertices, if any exist, in the second network. Our methodology identifies vertices in a local neighborhood of the vertices of interest in the first network that have verifiable corresponding vertices in the second network. 

Leveraging these known correspondences, referred to as seeds, we match the induced subgraphs in each network generated by the neighborhoods of these verified seeds, and rank the vertices of the second network in terms of the most likely matches to the original vertices of interest.


# Example 1: A Known Correspondence

Here, we produce a simple random graph (adjacency matrix) A with n =  and p = 0.5. We then produce B by maintaining the first m = 10 rows (the seeds), and permuting the remaining n - m rows.

```{r, fig.width = 5, fig.height=4}

# Number of seeds, total nubmer of vertices, and edge probability.
m <- 3
n <- 10
p <- 0.5
set.seed(123)
# Sample graph 1, and permute to graph 2; delete a vertex to show VN works
# when total number of vertices are different.
g1 = sample_sbm(n, as.matrix(p), n)
permute = c(1:m, sample(n-m)+m)
g2 = permute.vertices(g1, permute)
g2 = delete.vertices(g2, 10)

seeds <- matrix(cbind(1:m, 1:m), nrow = m)
x = 4
h = 1
ell = 2
R = 100
gamma = 0.01
vn = vnsgm(x,seeds,g1,g2,h,ell,R,gamma,plotF = TRUE)
```

Here green represents the set of seeds; red represents VOIs; blue represents unobserved vertices we matched via SGM. Shade of color represents how likely vertex i in graph 1 is a match to vertex j in graph 2.


# Example 2: r-Correlated SBM

Here, we sample (A, B) from an r-Correlated SBM via a random dot-product graph, permute the matrix like before, and observe the results of Seeded Graph Matching.

```{r, fig.width=4.5, fig.height=4}
# Define latent vectors for 2-SBM RDPG in X.
n <- 10
X1 <- matrix(rep(c(0.85, 0), n/2), nrow = n/2, byrow = TRUE)
X2 <- matrix(rep(c(0.3,0.8), n/2), nrow = n/2, byrow = TRUE)
X <- rbind(X1, X2)
set.seed(6789)

# Pearson correlation coefficient.
r <- 0.75

# Sample r-SBM.
sampled_graphs <- rdpg.sample.correlated(X, r)
g1 = sampled_graphs[[1]]
g2 = sampled_graphs[[2]]
A <- as_adj(g1, sparse = FALSE)
B <- as_adj(g2, sparse = FALSE)
# Display overlap.
gs.plot.plot_matrix(A + B, title="A + B (Overlap)", legend.name="A_ij + B_ij")
```

These graphs are not exactly equal, but have high correlation. Their matrix addition, pictured above, has many 2 and 0 entries, and few 1 entries. We now choose the first m/2 = 1 nodes from each block of each graph to be the seeds.

```{r, fig.width=4.5, fig.height=4}

# Identify seeds.
m <- 2
seed_indices <- c(1:(m/2), 1:(m/2)+n/2)
seeds <- matrix(c(seed_indices, seed_indices), nrow = m)

# Set up permutation.
block1_permutation <- sample(n/2 - m/2) + m/2
block2_permutation <- sample(n/2 - m/2) + m/2 + n/2
B_permutation <- c(1:(m/2), block1_permutation, 1:(m/2)+n/2, block2_permutation)
# Redefine B matrix and match using unordered sgm.
g2_p = permute.vertices(g2, B_permutation)
vn = vnsgm(x,seeds,g1,g2_p,h,ell,R,gamma,plotF = TRUE)
```

Here we display our original permutation for comparison purposes.
```{r, fig.width=5, fig.height=4}
rbind(c(1:10),B_permutation)
```







