---
title: "SIEM for Determination of Batch Effects in fMRI and dMRI Populations wrt Hemispheric Connectivity"
author: "Eric Bridgeford"
date: "December 31, 2017"
output: html_document
---

```{r setup, include=FALSE}
require(fmriutils)
require(graphstats)
require(mgc)
require(ggplot2)
require(latex2exp)
require(igraph)
require(stringr)
source('../../R/siem.R')

# accepts a matrix and thresholds/binarizes it
thresh_matrix = function(matrix, thresh=0.5) {
  thr = quantile(matrix, thresh)
  return(ifelse(matrix > thr, 1, 0))
}

# accepts a [n x n] adjacency matrix and computes the probabilities associated with an SBM
# where the vertices are grouped such that union_i(groups) = V(A) and
# intersection(group_i, group_j) = NULL for all i != j
block_data = function(matrix, groups) {
  # matrix is adwi_thresh n x n array
  # groups is a grouping of the vertices in the matrix as a list
  blocks = array(NaN, dim=c(2,2))
  for (i in 1:length(groups)) {
    for (j in 1:length(groups)) {
      blocks[i, j] = mean(matrix[groups[[i]], groups[[j]]])
    }
  }
  return(blocks)
}

parse_class <- function(basepath, dsets, subjects) {
  sex = list()
  # disease = list()
  age = list()
  include = c()
  for (dset in dsets) {
    path_to_file = file.path(basepath, paste(dset, "_phenotypic_data.csv", sep=""))
    tryCatch({
      tab = read.csv(path_to_file)
      tab$SEX[tab$SEX == '#' | is.na(tab$SEX) | is.nan(tab$SEX)] = NaN
      if (dset == 'KKI2009') {
        sexm <- 'M'
      } else if (dset == 'Templeton114') {
        sexm <- 1
      } else {
        sexm <- 2
      }
      tab$AGE_AT_SCAN_1 <- factor(tab$AGE_AT_SCAN_1)
      tab$SEX = tab$SEX == sexm
      tab = tab[complete.cases(tab$SEX),]
      tab$AGE_AT_SCAN_1 = as.numeric(levels(tab$AGE_AT_SCAN_1))[tab$AGE_AT_SCAN_1]
      for (idx in 1:dim(tab)[1]) {
        subid = toString(tab[idx,]$SUBID)
        sex[[subid]] = tab[idx,]$SEX
        age[[subid]] = tab[idx,]$AGE_AT_SCAN_1
        # disease[[subid]] = tab[idx,]$DSM_IV_TR
      }
    }, error=function(e) return(NaN))
  }
  
  sclass = array(NaN, dim=c(length(subjects)))
  ageclass = sclass
  # diseaseclass = sclass
  for (i in 1:length(subjects)) {
    subject = subjects[i]
    subid = sub('^0+(?=[1-9])', '', str_extract(subject, '(?<=sub-).*'), perl=TRUE)
    idx = which(names(sex) == subid)
    if (length(idx) >= 1) {
      sclass[i] <- tryCatch(sex[[subid]], error=NaN)
      ageclass[i] <- tryCatch(age[[subid]], error=NaN)
      # diseaseclass[i] <- disease[[subid]]
    }
  }
  return(list(sex=sclass, age=ageclass))#, disease=diseaseclass))
}
```

# Model

For details on the model, please see [SIEM model](http://docs.neurodata.io/graphstats/siem/siem.html). For an expanded look at the Homotopic vs. Heterotopic Connectivity Problem, see [Hemispheric Connectivity Notebook](http://docs.neurodata.io/graphstats/siem/hemispheric_within.html).

# Real Data Experiments

In this notebook, we determine whether there exists a batch effect in the difference in connectivity $p_{ipsi}$ ipsilaterally (same hemisphere) vs $p_{contra}$ contra-laterally (different region) connectivity within a particular modality. We consider $\delta_{x} = p_{ips} - p_{contra}$ to be the difference in connectivity for a graph from a graph or collection of graphs from a particular modality $x$.

## Tests

### for each Graph

Our test for this notebook is as follows:

\begin{align*}
  H_0: p_{ips} \leq p_{contra} \\
  H_A: p_{ips} > p_{contra}
\end{align*}

We will use a $1$-sample test to determine whether within a single graph we can determine a significant difference in connectivity.  

We will perform this experiment for both dMRI and fMRI-derived connectomes.

### for Determining a Batch

To determine a batch's test statistic, we will simply attempt to assess the magnitude of the differences in model fit between each pair of batches. To acquire a p-value, we will use a permutation-based approach. That is:

```
test_statistic(models, grouping):
  compute the average p_{homo} and p_{hetero} for each unique grouping label.
  For each i, j pair of unique grouping labels:
    stat[i, j] = |p_{homo, 1} - p_{homo, 2} - (p_{hetero, 1} - p_{hetero, 2})|
  return max(stats)  # test statistic is the pairing with the greatest magnitude
  
permutation_test(models, batch_grouping):
  compute the test statistic for the graphs given the default batch grouping.
  For i in 1:nrepetitions:
    permute the batch grouping to obtain a permuted grouping.
    compute the test statistic for the models given the permuted grouping.
  Compute the fraction of permuted test statistics > the given test statistic. 
```

Or the fraction of times that the maximum magnitude of a model fit difference exceeds the observed maximum magnitude of a model fit difference.

```{r}
case.experiment.within <- function(models, datasets.labs, split.labs, modal, nrep=1000) {
  # partition the models that do not have the split data
  include_sets <- which(!is.nan(split.labs))
  incl.models <- models[include_sets]
  incl.datasets.labs <- datasets.labs[include_sets]
  incl.split.labs <- split.labs[include_sets]
  
  dsets <- unique(incl.datasets.labs)
  results <- data.frame(dataset=c(), pval=c())
  
  for (i in 1:length(dsets)) {
    ss <- which(incl.datasets.labs == dsets[i])
    model_ss <- incl.models[ss]  # subset one dataset of models
    split_ss <- incl.split.labs[ss]
    if (length(unique(split_ss)) > 1) {
      tstat.alt <- gs.siem.batch.test(incl.models, split_ss, i=1, j=2)  # first edge group is homotopic, second is heterotopic
      tstat.nulls <- array(NaN, dim=c(nrep))
      for (j in 1:nrep) {
        permuted_ss <- sample(split_ss, size=length(split_ss))
        tstat.nulls[j] <-  gs.siem.batch.test(incl.models, permuted_ss, i=1, j=2)  # first edge group is homotopic, second is heterotopic
      }
      pval <- sum(tstat.nulls < tstat.alt)/nrep
    } else {
      pval <- NaN
    }
    results <- rbind(results, data.frame(dataset=dsets[i], pval=pval, modal=modal))
  }
  return(results)
}


case.experiment.between <- function(models, datasets.labs, modal, nrep=1000) {
  # partition the models that do not have the split data
  include_sets <- which(!is.nan(datasets.labs))
  models <- models[include_sets]
  datasets.labs <- datasets.labs[include_sets]
  
  dsets <- unique(datasets.labs)
  results <- data.frame(dataset=c(), pval=c())

  tstat.alt <- gs.siem.batch.test(models, datasets.labs, i=1, j=2)  # first edge group is homotopic, second is heterotopic
  tstat.nulls <- array(NaN, dim=c(nrep))
  for (j in 1:nrep) {
    permuted_ss <- sample(datasets.labs, size=length(datasets.labs))
    tstat.nulls[j] <-  gs.siem.batch.test(models, permuted_ss, i=1, j=2)  # first edge group is homotopic, second is heterotopic
  }
  results <- rbind(results, data.frame(dataset='set', pval=(sum(tstat.nulls > tstat.alt)/nrep), modal=modal))
  return(results)
}
```

## Raw Data

The data below can be downloaded and moved to appropriate folders as follows (note that the below section requires sudo access):

```
sudo mkdir /data/
sudo chmod -R 777 /data

cd /data
wget http://openconnecto.me/mrdata/share/derivatives/dwi_edgelists.tar.gz
wget http://openconnecto.me/mrdata/share/derivatives/fmri_edgelists.tar.gz
wget http://openconnecto.me/mrdata/share/connectome_stats/connectome_stats.zip

mkdir -p /data/connectome_stats /data/all_mr /data/all_mr/dwi/edgelists /data/all_mr/fmri/ranked/edgelists
mv dwi_edgelists.tar.gz /data/dwi/edgelists
cd /data/dwi/edgelists
tar -xvzf dwi_edgelists.tar.gz
mv /data/fmri_edgelists.tar.gz /data/fmri/ranked/edgelists
cd /data/fmri/ranked/edgelists
tar -xvzf fmri_edgelists.tar.gz
mv /data/connectome_stats.zip /data/connectome_stats.zip
cd /data/connectome_stats
unzip connectome_stats.zip
```

## Diffusion

```{r}
nroi <- 70
dwi.dsets = c('BNU1', 'BNU3', 'HNU1', 'KKI2009', 'NKI1', 'NKIENH', 'MRN1313', 'Templeton114', 'Templeton255', 'SWU4')
dwi.atlas = 'desikan'
dwi.basepath = '/data/all_mr/dwi/edgelists'

graphobj = fmriu.io.collection.open_graphs(basepath = dwi.basepath, atlases = dwi.atlas, datasets = dwi.dsets,
                                           gname = 'graphs', fmt='edgelist', rtype = 'array')
dwi.graphs = graphobj$graphs
dwi.datasets = graphobj$dataset
dwi.subjects = graphobj$subjects
dwi.sessions = graphobj$sessions

sexpath = '/data/all_mr/phenotypic/'
class = parse_class(sexpath, dwi.dsets, dwi.subjects)
dwi.sexs = class$sex
```

```{r}
ne = 1225
nroi <- 70
group1 <- c()  # edges in same hemisphere
group2 <- c()  # edges across hemispheres
for (i in 1:nroi) {
  for (j in 1:nroi) {
    idx <- (i - 1)*nroi + j
    if ((i <= 35 & j <= 35) | (i > 35 & j > 35)) {
      group1 <- c(group1, idx)
    } else {
      group2 <- c(group2, idx)
    }
  }
}
Es <- list(group1, group2)
dwi.models <- sapply(1:dim(dwi.graphs)[1], function(i) {
                  gs.siem.fit(thresh_matrix(dwi.graphs[i,,], 0), Es, alt='greater')
                }, simplify = FALSE)
```


# Functional

```{r}
nroi <- 70
fmri.dsets = c('BNU1', 'BNU2', 'BNU3', 'HNU1', 'IBATRT', 'IPCAS1', 'IPCAS2', 'IPCAS5', 'IPCAS6', 'IPCAS8', 'MRN1', 'NYU1', 'SWU1', 'SWU2', 'SWU3', 'SWU4', 'UWM', 'XHCUMS')
fmri.atlas = 'desikan-2mm'
fmri.basepath = '/data/all_mr/fmri/ranked/edgelists/'

graphobj = fmriu.io.collection.open_graphs(basepath = fmri.basepath, atlases = fmri.atlas, datasets=fmri.dsets, fmt='edgelist', rtype = 'array')
fmri.graphs = graphobj$graphs
fmri.datasets = graphobj$dataset
fmri.subjects = graphobj$subjects
fmri.sessions <- graphobj$sessions

sexpath = '/data/all_mr/phenotypic/'
class = parse_class(sexpath, fmri.dsets, fmri.subjects)
fmri.sexs = class$sex
```

```{r}
ne = 1225
nroi <- 70
group1 <- c()  # edges in same hemisphere
group2 <- c()  # edges across hemispheres
for (i in 1:nroi) {
  for (j in 1:nroi) {
    idx <- (i - 1)*nroi + j
    if ((i <= 35 & j <= 35) | (i > 35 & j > 35)) {
      group1 <- c(group1, idx)
    } else {
      group2 <- c(group2, idx)
    }
  }
}
Es <- list(group1, group2)
fmri.models <- sapply(1:dim(fmri.graphs)[1], function(i) {
                  gs.siem.fit(thresh_matrix(fmri.graphs[i,,], 0.5), Es, alt='greater')
                }, simplify = FALSE)
```

# Case 1: Session Partitioning

+ Is there a batch effect between sessions of a particular study?

## Procedure
 
1) Compute test statistic given the default partitioning, and obtain $\tau_{observed}$, as the maximum magnitude of difference in homotopic vs. heterotopic connectivity between any pair of 2 sessions.
2) permute the set labels of the combined set $nperm$ times (maintaining sex of each graph) to obtain the distribution of $\tau_{null}$, reporting $\hat{p}$, the estimator of $\mathbb{E}\left[\tau_{null} < \tau_{observed}\right]$

```{r}
results <- rbind(case.experiment.within(dwi.models, dwi.datasets, dwi.sessions, 'dMRI', nrep=1000),
      case.experiment.within(fmri.models, fmri.datasets, fmri.sessions, 'fMRI', nrep=1000))
```

```{r}
ggplot(results, aes(x=modal, y=pval, group=dataset, color=dataset)) +
  geom_point(size=2) +
  xlab("Modality") +
  ylab(TeX("$p$-Value")) +
  ggtitle("Case 1: Session Partitioning")
```

# Case 2: Subject Partitioning

+ Is there a batch effect between subjects in a particular study?

## Procedure
 
1) Compute test statistic given the default partitioning, and obtain $\tau_{observed}$, as the maximum magnitude of difference in homotopic vs. heterotopic connectivity between any pair of 2 subjects.
2) permute the set labels of the combined set $nperm$ times (maintaining sex of each graph) to obtain the distribution of $\tau_{null}$, reporting $\hat{p}$, the estimator of $\mathbb{E}\left[\tau_{null} < \tau_{observed}\right]$

```{r}
results <- rbind(case.experiment.within(dwi.models, dwi.datasets, dwi.subjects, 'dMRI', nrep=1000),
      case.experiment.within(fmri.models, fmri.datasets, fmri.subjects, 'fMRI', nrep=1000))
```

```{r}
ggplot(results, aes(x=modal, y=pval, group=dataset, color=dataset)) +
  geom_point(size=2) +
  xlab("Modality") +
  ylab(TeX("$p$-Value")) +
  ggtitle("Case 2: Subject Partitioning")
```

# Case 3: Same Sex

+ Is there a batch effect between sexs in a particular study?

## Procedure
 
1) Compute test statistic given the default partitioning, and obtain $\tau_{observed}$, as the maximum magnitude of difference in homotopic vs. heterotopic connectivity between any pair of 2 subjects.
2) permute the set labels of the combined set $nperm$ times (maintaining sex of each graph) to obtain the distribution of $\tau_{null}$, reporting $\hat{p}$, the estimator of $\mathbb{E}\left[\tau_{null} < \tau_{observed}\right]$

```{r}
results <- rbind(case.experiment.within(dwi.models, dwi.datasets, dwi.sexs, 'dMRI', nrep=1000),
      case.experiment.within(fmri.models, fmri.datasets, fmri.sexs, 'fMRI', nrep=1000))
```

```{r}
ggplot(results, aes(x=modal, y=pval, group=dataset, color=dataset)) +
  geom_point(size=2) +
  xlab("Modality") +
  ylab(TeX("$p$-Value")) +
  ggtitle("Case 3: Sex Partitioning")
```

# Case 4: Same Site

+ Is there a batch effect between studies with site held fixed?  

## Procedure
 
1) Compute test statistic given the default partitioning, and obtain $\tau_{observed}$, as the maximum magnitude of difference in homotopic vs. heterotopic connectivity between any pair of 2 subjects.
2) permute the set labels of the combined set $nperm$ times (maintaining sex of each graph) to obtain the distribution of $\tau_{null}$, reporting $\hat{p}$, the estimator of $\mathbb{E}\left[\tau_{null} < \tau_{observed}\right]$

```{r}
dwi.studies <- dwi.datasets
fmri.studies <- fmri.datasets
dwi.studies[dwi.studies == 'BNU1' | dwi.studies == 'BNU3'] <- 'BNU'
fmri.studies[fmri.studies == 'BNU1' | fmri.studies == 'BNU3'] <- 'BNU'
dwi.studies[dwi.studies == 'Templeton114' | dwi.studies == 'Templeton255'] <- 'Templeton'

results <- rbind(case.experiment.within(dwi.models, dwi.studies, dwi.datasets, 'dMRI', nrep=1000),
      case.experiment.within(fmri.models, fmri.studies, fmri.datasets, 'fMRI', nrep=1000))
```

```{r}
ggplot(results, aes(x=modal, y=pval, group=dataset, color=dataset)) +
  geom_point(size=2) +
  xlab("Modality") +
  ylab(TeX("$p$-Value")) +
  ggtitle("Case 4: Same Site")
```

# Case 5: Same Demographics

+ Is there a batch effect between sites with demographic held fixed?  

## Procedure
 
1) Compute test statistic given the default partitioning, and obtain $\tau_{observed}$, as the maximum magnitude of difference in homotopic vs. heterotopic connectivity between any pair of 2 subjects.
2) permute the set labels of the combined set $nperm$ times (maintaining sex of each graph) to obtain the distribution of $\tau_{null}$, reporting $\hat{p}$, the estimator of $\mathbb{E}\left[\tau_{null} < \tau_{observed}\right]$

```{r}
dwi.dset.sset <- which(dwi.datasets %in% c('BNU1', 'SWU4', 'HNU1', 'BNU3'))
dwi.models.subs  <- dwi.models[dwi.dset.sset]; dwi.datasets.subs <- dwi.datasets[dwi.dset.sset]

fmri.dset.sset <- which(fmri.datasets %in% c('BNU1', 'SWU4', 'HNU1', 'BNU3'))
fmri.models.subs  <- fmri.models[fmri.dset.sset]; fmri.datasets.subs <- fmri.datasets[fmri.dset.sset]

results <- rbind(case.experiment.between(dwi.models.subs, dwi.datasets.subs, 'dMRI', nrep=1000),
      case.experiment.between(fmri.models.subs, fmri.datasets.subs, 'fMRI', nrep=1000))
```

```{r}
ggplot(results, aes(x=modal, y=pval)) +
  geom_point(size=2) +
  xlab("Modality") +
  ylab(TeX("$p$-Value")) +
  ggtitle("Case 5: Same Demographics")
```

# Case 6: Disparate Demographics

+  Is there a batch effect between sites without demographic held fixed?  

## Procedure
 
1) Compute test statistic given the default partitioning, and obtain $\tau_{observed}$, as the maximum magnitude of difference in homotopic vs. heterotopic connectivity between any pair of 2 subjects.
2) permute the set labels of the combined set $nperm$ times (maintaining sex of each graph) to obtain the distribution of $\tau_{null}$, reporting $\hat{p}$, the estimator of $\mathbb{E}\left[\tau_{null} < \tau_{observed}\right]$

```{r}
results <- rbind(case.experiment.between(dwi.models, dwi.datasets, 'dMRI', nrep=1000),
      case.experiment.between(fmri.models, fmri.datasets, 'fMRI', nrep=1000))
```

```{r}
ggplot(results, aes(x=modal, y=pval)) +
  geom_point(size=2) +
  xlab("Modality") +
  ylab(TeX("$p$-Value")) +
  ggtitle("Case 6: Disparate Demographics")
```
