---
title: "Connectome Coding Notebook"
output:
  html_document:
    df_print: paged
---

# Exploratory Plots

```{r}
require(igraph)
require(graphstats)
require(gridExtra)
require(ggplot2)
require(gridExtra)
require(ggpubr)
require(reshape2)
require(grid)

nan.sum <- function(x) {sum(x, na.rm=TRUE)}

g_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)
}
#' Graph Heatmap plot
#'
#' A function that plots an igraph object, as a heatmap.
#'
#' @import ggplot2
#' @import igraph
#' @importFrom reshape2 melt
#' @importFrom ggpubr as_ggplot
#' @param g input graph, as an igraph object. See \code{\link[igraph]{graph}} for details.
#' @param title the title for the square plot. Defaults to \code{""}.
#' @param src.label the source label for the graph. Defaults to \code{"Vertex"}.
#' @param tgt.label the target label for the graph. Defaults to \code{"Vertex"}.
#' @param edge.attr the name of the attribute to use for weights. Defaults to \code{NULL}.
#' \itemize{
#' \item{\code{is.null(edge.attr)} plots the graph as a binary adjacency matrix.}
#' \item{\code{is.character(edge.attr)} plot the graph as a weighted adjacency matrix, with edge-weights for \code{E(g)} given by \code{E(g)[[edge.attr]]}.}
#' }
#' @param font.size the default font size for the plot text. Axis/legend text is \code{font.size - 2}. Defaults to \code{NULL}.
#' \itemize{
#' \item{\code{is.null(font.size)} uses the default sizing for all fonts.}
#' \item{\code{!is.null(font.size)} uses \code{font.size} as the font sizing for the plots.}
#' }
#' @param vertex.label an attribute for naming the vertices. Defaults to \code{NULL}.
#' \itemize{
#' \item{\code{vertex.label==FALSE} name the vertices \code{V(g)} sequentially, as 1, 2, ... n.}
#' \item{\code{vertex.label==TRUE} name the vertices \code{V(g)} as \code{V(g)$name}.}
#' }
#' @param vertex.attr an attribute to color vertices. Defaults to \code{FALSE}.
#' \itemize{
#' \item{\code{vertex.attr==FALSE} assumes no grouping of the vertices, and adds no color accordingly.}
#' \item{\code{is.character(vertex.attr)} assumes a grouping of the vertices given for \code{V(g)} by \code{V(G)[[vertex.attr]]}, and groups the vertices in \code{V(g)} into ordered blocks with color-coding.}
#' }
#' @param edge.xfm log-transform the edge-weights. Defaults to \code{FALSE}.
#' \itemize{
#' \item{\code{edge.xfm==FALSE} do not transform the edge-weights.}
#' \item{\code{edge.xfm == "log"} transform the edge values, using the natural-logarithm operation. See \code{\link[base]{log}} for details. Does not work if there are negative edge-weights.}
#' \item{\code{edge.xfm == "log10"} transform the edge values, using the logarithm-base-10 operation. See \code{\link[base]{log10}} for details. Does not work if there are negative edge-weights.}
#' \item{\code{! (edge.xfm %in% c(FALSE, "log", "log10"))} assumes \code{edge.xfm} is a function, and transform the vertices with the provided function.}
#' }
#' @return the graph/graphs as a plot.
#' @author Eric Bridgeford
#' @export
gs.plot.heatmap <- function(g, title="",src.label="Vertex", tgt.label="Vertex", edge.attr=NULL,
                            font.size=NULL, vertex.label=FALSE, vertex.attr=FALSE, edge.xfm=FALSE, eps=0.0001,
                            degree=TRUE) {
  # load adjacency matrix as a dense matrix
  adj <- as_adjacency_matrix(g, attr=edge.attr, names=vertex.label, type="both", sparse=FALSE)
  adj.data <- melt(adj)  # melt the graph to a data-frame with row and colnames preserved
  colnames(adj.data) <- c("Source", "Target", "Weight")
  alpha = 1
  hist.src <- apply(adj, c(1), nan.sum)
  hist.tgt <- apply(adj, c(2), nan.sum)
  edge.colors=c("#020202")
  if (!is.character(edge.attr)) {
    adj.data$Weight <- factor(adj.data$Weight, levels=c(0, 1), ordered=TRUE)
    wt.name <- "Connection"
  } else {
    wt.name <- edge.attr
    if (edge.xfm != FALSE) {
      if (edge.xfm == "log") {
        edge.xfm=log
        # set 0-weight edges to far lower than rest of graph
        adj.data$Weight <- adj.data$Weight + eps
        wt.name = sprintf("log(%s)", wt.name)
      } else if (edge.xfm == "log10") {
        edge.xfm = log10
        # set 0-weight edges to far lower than rest of graph
        adj.data$Weight <- adj.data$Weight + eps
        wt.name = sprintf("log10(%s)", wt.name)
      }
      adj.data$Weight <- do.call(edge.xfm, list(adj.data$Weight))
    }
  }
  hist.dat <- rbind(data.frame(Vertex=1:dim(adj)[1], Degree=hist.src/max(hist.src), Type=wt.name, direction="Source"),
                    data.frame(Vertex=1:dim(adj)[1], Degree=hist.tgt/max(hist.tgt), Type=wt.name, direction="Target"))

  thm = list(theme_void(),
               guides(fill=FALSE),
               theme(plot.margin=unit(rep(0,4), "lines")))
  plot.adj <- ggplot(adj.data, aes(x=Source, y=Target, alpha=Weight)) +
    geom_tile(fill=edge.colors) +
    xlab(src.label) +
    ylab(tgt.label) +
    ggtitle(title) +
    theme_bw() +
    theme(rect=element_blank(), panel.grid=element_blank()) +
    thm
  if (vertex.label) {
    plot.adj <- plot.adj + theme(axis.text.x = element_text(angle=60, hjust=1))
  }
  if (is.character(edge.attr)) {
    plot.adj <- plot.adj +
      scale_fill_gradientn(colors=edge.colors, name=wt.name) +
      guides(alpha=guide_legend(title=wt.name))
  } else {
    plot.adj <- plot.adj +
      scale_fill_manual(values=edge.colors, name=wt.name) +
      guides(alpha=guide_legend(title=wt.name))
  }
  if (degree) {
    thm = list(theme_void(),
               guides(fill=FALSE),
               theme(plot.margin=unit(rep(0,4), "lines"), legend.position=NaN))
    top.plot <- ggplot(subset(hist.dat, direction == "Source"), aes(x=Vertex, y=Degree, fill=Type, group=Type)) +
      geom_bar(stat = "identity", position="identity", alpha=alpha) +
      scale_fill_manual(values=edge.colors) +
      thm
    right.plot <- ggplot(subset(hist.dat, direction == "Target"), aes(x=Vertex, y=Degree, fill=Type, group=Type)) +
      geom_bar(stat = "identity", position="identity", alpha=alpha) +
      scale_fill_manual(values=edge.colors) +
      coord_flip() +
      thm
    empty <- ggplot() + geom_blank() + thm
    pleg <- g_legend(plot.adj)
    widths=c(0.6, 0.2, 0.2)
    heights=c(0.2, 0.8)
    plot.adj <- as_ggplot(arrangeGrob(grobs=list(top.plot, empty, empty, plot.adj + theme(legend.position=NaN, title=element_blank()), right.plot, pleg), byrow=TRUE,
                                      ncol=3, nrow=2, widths=widths, heights=heights, top=textGrob(title, gp=gpar(cex=1.3)),
                                      left=tgt.label, bottom=src.label))
  }
  #if (is.character(vertex.attr)) {
  #  vertices <- colnames(adj)
  #  jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))
  #  V.gr <- factor(get.vertex.attribute(g, vertex.attr))  # group the vertices by their attribute
  #  un.vertices <- levels(V.gr)
  #  attr.colors <- jet.colors(length(un.vertices))
  #  for (i in 1:length(attr.colors)) {
  #    vertex.num <- min(which())
  #    plot.adj <- plot.adj +
  #      geom_rect()
  #  }
  #  plot.adj <- plot.adj +
  #    geom_tile(data=adj.data, aes(x=Source, y=Target, fill=Fill1), show.legend=TRUE, alpha=0.2) +
  #    geom_tile(data=adj.data, aes(x=Source, y=Target, fill=Fill2), show.legend=FALSE, alpha=0.2)
  #}
  return(plot.adj)
}

#' Graph Grid plot
#'
#' A function that plots an igraph object, as a grid, with intensity denoted by the size of dots on the grid.
#'
#' @import ggplot2
#' @import igraph
#' @importFrom reshape2 melt
#' @importFrom gridExtra arrangeGrob
#' @importFrom ggpubr as_ggplot
#' @param g input graph, as an igraph object. See \code{\link[igraph]{graph}} for details.
#' @param title the title for the square plot. Defaults to \code{""}.
#' @param src.label the source label for the graph. Defaults to \code{"Vertex"}.
#' @param tgt.label the target label for the graph. Defaults to \code{"Vertex"}.
#' @param edge.attr the name of the attribute to use for weights. Defaults to \code{NULL}. Can be a list of `edge.attr` if you want to overlay different edge-attributes on the same plot. Supports up to 4 edge-attributes at once.
#' \itemize{
#' \item{\code{is.null(edge.attr)} plots the graph as a binary adjacency matrix.}
#' \item{\code{is.character(edge.attr)} plot the graph as a weighted adjacency matrix, with edge-weights for \code{E(g)} given by \code{E(g)[[edge.attr]]}.}
#' }
#' @param font.size the default font size for the plot text. Axis/legend text is \code{font.size - 2}. Defaults to \code{NULL}.
#' \itemize{
#' \item{\code{is.null(font.size)} uses the default sizing for all fonts.}
#' \item{\code{!is.null(font.size)} uses \code{font.size} as the font sizing for the plots.}
#' }
#' @param vertex.label an attribute for naming the vertices. Defaults to \code{NULL}.
#' \itemize{
#' \item{\code{vertex.label==FALSE} name the vertices \code{V(g)} sequentially, as 1, 2, ... n.}
#' \item{\code{vertex.label==TRUE} name the vertices \code{V(g)} as \code{V(g)$name}.}
#' }
#' @param vertex.attr an attribute to color vertices. Defaults to \code{FALSE}.
#' \itemize{
#' \item{\code{vertex.attr==FALSE} assumes no grouping of the vertices, and adds no color accordingly.}
#' \item{\code{is.character(vertex.attr)} assumes a grouping of the vertices given for \code{V(g)} by \code{V(G)[[vertex.attr]]}, and groups the vertices in \code{V(g)} into ordered blocks with color-coding.}
#' }
#' @param edge.xfm transform the edge-weights. Defaults to \code{FALSE}. Can be a list of `edge.xfm` if you want to overlay different edge-attributes on the same plot in the plot with different transforms for each.
#' \itemize{
#' \item{\code{edge.xfm==FALSE} do not transform the edge-weights.}
#' \item{\code{edge.xfm == "log"} transform the edge values, using the natural-logarithm operation. See \code{\link[base]{log}} for details. Does not work if there are negative edge-weights. Pads with `eps << min(edge-weight)` if there are entries of zero in the graph.}
#' \item{\code{edge.xfm == "log10"} transform the edge values, using the logarithm-base-10 operation. See \code{\link[base]{log10}} for details. Does not work if there are negative edge-weights. Pads with `eps << min(edge-weight)` if there are entries of zero in the graph.}
#' \item{\code{! (edge.xfm %in% c(FALSE, "log", "log10"))} assumes \code{edge.xfm} is a function, and transform the vertices with the provided function.}
#' }
#' @param eps if you specify an `edge.xfm` that is logarithmic, indicate the padding that zero entries should receive. Defaults to `.0001`. `eps` should be `<< min(edge-weight)`.
#' @param degree Whether to plot the marginal vertex degrees. Defaults to `FALSE`.
#' @return the graph/graphs as a plot.
#' @author Eric Bridgeford
#' @export
gs.plot.grid <- function(g, title="",src.label="Vertex", tgt.label="Vertex", edge.attr=NULL,
                         font.size=NULL, vertex.label=FALSE, vertex.attr=FALSE, edge.xfm=FALSE, eps=0.0001,
                         degree=FALSE) {
  if (!is.vector(edge.attr)) {
    edge.attr <- list(edge.attr)
  }
  if (length(edge.xfm) < length(edge.attr)) {
    edge.xfm <- rep(edge.xfm[1], min(length(edge.attr)))
  }
  cvec <- c("#020202", "#f41711", "#94d6c9", "#5f8793")
  cvec <- cvec[1:length(edge.attr)]
  names(cvec) <- edge.attr
  alpha <- 1/length(edge.attr)
  plot.dat <- data.frame()
  hist.dat <- data.frame()
  for (i in 1:length(edge.attr)) {
    attr <- edge.attr[[i]]; xfm <- edge.xfm[i]
    adj <- as_adjacency_matrix(g, attr=attr, names=vertex.label, type="both", sparse=FALSE)
    adj.data <- melt(adj)  # melt the graph to a data-frame with row and colnames preserved
    colnames(adj.data) <- c("Source", "Target", "Weight")
    hist.src <- apply(adj, c(1), nan.sum)
    hist.tgt <- apply(adj, c(2), nan.sum)
    if (is.null(attr)) {
       wt.name = "Connection"
    } else {
      wt.name <- attr
    }
    if (xfm != FALSE) {
      if (xfm == "log") {
        xfm=log
        # set 0-weight edges to far lower than rest of graph
        adj.data$Weight <- adj.data$Weight + eps
        wt.name = sprintf("log(%s)", wt.name)
      } else if (edge.xfm == "log10") {
        xfm = log10
        # set 0-weight edges to far lower than rest of graph
        adj.data$Weight <- adj.data$Weight + eps
        wt.name = sprintf("log10(%s)", wt.name)
      }
      adj.data$Weight <- do.call(xfm, list(adj.data$Weight))
    }
    adj.data <- adj.data[adj.data$Weight != 0,]
    if (length(edge.attr) > 1) {
      adj.data$Weight <- (adj.data$Weight - min(adj.data$Weight, na.rm=TRUE))/(max(adj.data$Weight, na.rm=TRUE) - min(adj.data$Weight, na.rm=TRUE))  # normalize on 0-1
    }
    plot.dat <- rbind(plot.dat, cbind(adj.data, Type=wt.name))
    hist.dat <- rbind(hist.dat, rbind(data.frame(Vertex=1:dim(adj)[1], Degree=hist.src/max(hist.src), Type=wt.name, direction="Source"),
                                      data.frame(Vertex=1:dim(adj)[1], Degree=hist.tgt/max(hist.tgt), Type=wt.name, direction="Target")))
  }
  hist.dat$Degree <- as.vector(hist.dat$Degree)
  thm = list(theme_void(),
               guides(fill=FALSE),
               theme(plot.margin=unit(rep(0,4), "lines")))
  plot.adj <- ggplot(plot.dat, aes(x=Source, y=Target, size=Weight, color=Type, group=Type)) +
    geom_point(alpha=alpha) +
    xlab(src.label) +
    ylab(tgt.label) +
    ggtitle(title) +
    theme_bw() +
    theme(rect=element_blank(), panel.grid=element_blank()) +
    guides(color=guide_legend(order=2), size=guide_legend(order=1), alpha=FALSE) +
    thm
  plot.adj <- plot.adj + scale_color_manual(values=cvec)
  if (vertex.label) {
    plot.adj <- plot.adj + theme(axis.text.x = element_text(angle=60, hjust=1))
  } else {
    plot.adj <- plot.adj +
      xlim(1, dim(adj)[1]) +
      ylim(1, dim(adj)[1])
  }
  if (degree) {
    thm = list(theme_void(),
               guides(fill=FALSE),
               theme(plot.margin=unit(rep(0,4), "lines"), legend.position=NaN))
    top.plot <- ggplot(subset(hist.dat, direction == "Source"), aes(x=Vertex, y=Degree, fill=Type, group=Type)) +
      geom_bar(stat = "identity", position="identity", alpha=alpha) +
      scale_fill_manual(values=cvec) +
      thm
    right.plot <-  ggplot(subset(hist.dat, direction == "Target"), aes(x=Vertex, y=Degree, fill=Type, group=Type)) +
      geom_bar(stat = "identity", position="identity", alpha=alpha) +
      scale_fill_manual(values=cvec) +
      coord_flip() +
      thm
    empty <- ggplot() + geom_blank() + thm
    pleg <- g_legend(plot.adj)
    widths=c(0.6, 0.2, 0.2)
    heights=c(0.2, 0.8)
    plot.adj <- as_ggplot(arrangeGrob(grobs=list(top.plot, empty, empty, plot.adj + theme(legend.position=NaN, title=element_blank()), right.plot, pleg), byrow=TRUE,
                                      ncol=3, nrow=2, widths=widths, heights=heights, top=textGrob(title, gp=gpar(cex=1.3)), left=tgt.label,
                                      bottom=src.label))
  }
  #if (is.character(vertex.attr)) {
  # reorder the vertices so that vertices in same group are together

  #}
  #if (is.character(vertex.attr)) {
  #  vertices <- colnames(adj)
  #  jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))
  #  V.gr <- factor(get.vertex.attribute(g, vertex.attr))  # group the vertices by their attribute
  #  un.vertices <- levels(V.gr)
  #  attr.colors <- jet.colors(length(un.vertices))
  #  for (i in 1:length(attr.colors)) {
  #    vertex.num <- min(which())
  #    plot.adj <- plot.adj +
  #      geom_rect()
  #  }
  #  plot.adj <- plot.adj +
  #    geom_tile(data=adj.data, aes(x=Source, y=Target, fill=Fill1), show.legend=TRUE, alpha=0.2) +
  #    geom_tile(data=adj.data, aes(x=Source, y=Target, fill=Fill2), show.legend=FALSE, alpha=0.2)
  #}
  return(plot.adj)
}

#' Graph Grid plot
#'
#' A function that plots an igraph object, as a series of overlapping heatmaps, with intensity denoted by the colorscale of dots on the heatmap.
#'
#' @import ggplot2
#' @import igraph
#' @importFrom reshape2 melt
#' @importFrom gridExtra arrangeGrob
#' @importFrom ggpubr as_ggplot
#' @param g input graph, as an igraph object. See \code{\link[igraph]{graph}} for details.
#' @param title the title for the square plot. Defaults to \code{""}.
#' @param src.label the source label for the graph. Defaults to \code{"Vertex"}.
#' @param tgt.label the target label for the graph. Defaults to \code{"Vertex"}.
#' @param edge.attr the name of the attribute to use for weights. Defaults to \code{NULL}. Can be a list of `edge.attr` if you want to overlay different edge-attributes on the same plot. Supports up to 4 edge-attributes at once.
#' \itemize{
#' \item{\code{is.null(edge.attr)} plots the graph as a binary adjacency matrix.}
#' \item{\code{is.character(edge.attr)} plot the graph as a weighted adjacency matrix, with edge-weights for \code{E(g)} given by \code{E(g)[[edge.attr]]}.}
#' }
#' @param font.size the default font size for the plot text. Axis/legend text is \code{font.size - 2}. Defaults to \code{NULL}.
#' \itemize{
#' \item{\code{is.null(font.size)} uses the default sizing for all fonts.}
#' \item{\code{!is.null(font.size)} uses \code{font.size} as the font sizing for the plots.}
#' }
#' @param vertex.label an attribute for naming the vertices. Defaults to \code{NULL}.
#' \itemize{
#' \item{\code{vertex.label==FALSE} name the vertices \code{V(g)} sequentially, as 1, 2, ... n.}
#' \item{\code{vertex.label==TRUE} name the vertices \code{V(g)} as \code{V(g)$name}.}
#' }
#' @param vertex.attr an attribute to color vertices. Defaults to \code{FALSE}.
#' \itemize{
#' \item{\code{vertex.attr==FALSE} assumes no grouping of the vertices, and adds no color accordingly.}
#' \item{\code{is.character(vertex.attr)} assumes a grouping of the vertices given for \code{V(g)} by \code{V(G)[[vertex.attr]]}, and groups the vertices in \code{V(g)} into ordered blocks with color-coding.}
#' }
#' @param edge.xfm transform the edge-weights. Defaults to \code{FALSE}. Can be a list of `edge.xfm` if you want to overlay different edge-attributes on the same plot in the plot with different transforms for each.
#' \itemize{
#' \item{\code{edge.xfm==FALSE} do not transform the edge-weights.}
#' \item{\code{edge.xfm == "log"} transform the edge values, using the natural-logarithm operation. See \code{\link[base]{log}} for details. Does not work if there are negative edge-weights. Pads with `eps << min(edge-weight)` if there are entries of zero in the graph.}
#' \item{\code{edge.xfm == "log10"} transform the edge values, using the logarithm-base-10 operation. See \code{\link[base]{log10}} for details. Does not work if there are negative edge-weights. Pads with `eps << min(edge-weight)` if there are entries of zero in the graph.}
#' \item{\code{! (edge.xfm %in% c(FALSE, "log", "log10"))} assumes \code{edge.xfm} is a function, and transform the vertices with the provided function.}
#' }
#' @param eps if you specify an `edge.xfm` that is logarithmic, indicate the padding that zero entries should receive. Defaults to `.0001`. `eps` should be `<< min(edge-weight)`.
#' @param degree Whether to plot the marginal vertex degrees. Defaults to `FALSE`.
#' @return the graph/graphs as a plot.
#' @author Eric Bridgeford
#' @export
gs.plot.heatmap_overlay <- function(g, title="", src.label="Vertex", tgt.label="Vertex", edge.attr=NULL,
                                    font.size=NULL, vertex.label=FALSE, vertex.attr=FALSE, edge.xfm=FALSE, eps=0.0001,
                                    degree=FALSE) {
  if (!is.vector(edge.attr)) {
    edge.attr <- list(edge.attr)
  }
  if (length(edge.xfm) < length(edge.attr)) {
    edge.xfm <- rep(edge.xfm[1], min(length(edge.attr)))
  }
  cvec <- c("#020202", "#f41711", "#94d6c9", "#5f8793")
  cvec <- cvec[1:length(edge.attr)]
  names(cvec) <- edge.attr
  alpha <- 1/(length(edge.attr))
  plot.dat <- data.frame()
  hist.dat <- data.frame()
  for (i in 1:length(edge.attr)) {
    attr <- edge.attr[[i]]; xfm <- edge.xfm[i]
    adj <- as_adjacency_matrix(g, attr=attr, names=vertex.label, type="both", sparse=FALSE)
    adj.data <- melt(adj)  # melt the graph to a data-frame with row and colnames preserved
    colnames(adj.data) <- c("Source", "Target", "Weight")
    hist.src <- apply(adj, c(1), nan.sum)
    hist.tgt <- apply(adj, c(2), nan.sum)
    if (is.null(attr)) {
      wt.name = "Connection"
    } else {
      wt.name <- attr
    }
    if (xfm != FALSE) {
      if (xfm == "log") {
        xfm=log
        # set 0-weight edges to far lower than rest of graph
        adj.data$Weight <- adj.data$Weight + eps
        wt.name = sprintf("log(%s)", wt.name)
      } else if (edge.xfm == "log10") {
        xfm = log10
        # set 0-weight edges to far lower than rest of graph
        adj.data$Weight <- adj.data$Weight + eps
        wt.name = sprintf("log10(%s)", wt.name)
      }
      adj.data$Weight <- do.call(xfm, list(adj.data$Weight))
    }
    adj.data <- adj.data[adj.data$Weight != 0,]
    if (length(edge.attr) > 1) {
      adj.data$Weight <- (adj.data$Weight - min(adj.data$Weight, na.rm=TRUE))/(max(adj.data$Weight, na.rm=TRUE) - min(adj.data$Weight, na.rm=TRUE))  # normalize on 0-1
    }
    plot.dat <- rbind(plot.dat, cbind(adj.data, Type=wt.name))
    hist.dat <- rbind(hist.dat, rbind(data.frame(Vertex=1:dim(adj)[1], Degree=hist.src/max(hist.src), Type=wt.name, direction="Source"),
                                      data.frame(Vertex=1:dim(adj)[1], Degree=hist.tgt/max(hist.tgt), Type=wt.name, direction="Target")))
  }
  hist.dat$Degree <- as.vector(hist.dat$Degree)

  thm = list(theme_void(),
               guides(fill=FALSE),
               theme(plot.margin=unit(rep(0,4), "lines")))
  plot.adj <- ggplot(subset(plot.dat, Type == edge.attr[1]), aes(x=Source, y=Target, fill=Type, alpha=Weight, color=Type, group=Type), alpha=alpha) +
    geom_tile() +
    xlab(src.label) +
    ylab(tgt.label) +
    ggtitle(title) +
    theme_bw() +
    theme(rect=element_blank(), panel.grid=element_blank()) +
    thm
  plot.adj <- plot.adj + scale_color_manual(values=cvec) +
    scale_fill_manual(values=cvec)

  if (length(edge.attr) > 1) {
    for (i in 2:length(edge.attr)) {
      plot.adj <- plot.adj +
        geom_point(data=subset(plot.dat, Type == edge.attr[i]), aes(x=Source, y=Target, size=Weight, fill=Type, color=Type, group=Type), alpha=1)
    }
    plot.adj <- plot.adj +
      scale_size_continuous(range = c(0,2))
  }
  if (vertex.label) {
    plot.adj <- plot.adj + theme(axis.text.x = element_text(angle=60, hjust=1))
  }
  if (degree) {
    thm = list(theme_void(),
               guides(fill=FALSE),
               theme(plot.margin=unit(rep(0,4), "lines"), legend.position=NaN))
    top.plot <- ggplot(subset(hist.dat, direction == "Source"), aes(x=Vertex, y=Degree, fill=Type, group=Type)) +
      geom_bar(stat = "identity", position="identity", alpha=alpha) +
      scale_fill_manual(values=cvec) +
      thm
    right.plot <-  ggplot(subset(hist.dat, direction == "Target"), aes(x=Vertex, y=Degree, fill=Type, group=Type)) +
      geom_bar(stat = "identity", position="identity", alpha=alpha) +
      scale_fill_manual(values=cvec) +
      coord_flip() +
      thm
    empty <- ggplot() + geom_blank() + thm
    pleg <- g_legend(plot.adj)
    widths=c(0.6, 0.2, 0.2)
    heights=c(0.2, 0.8)
    plot.adj <- as_ggplot(arrangeGrob(grobs=list(top.plot, empty, empty, plot.adj + theme(legend.position=NaN, title=element_blank()), right.plot, pleg), byrow=TRUE,
                                      ncol=3, nrow=2, widths=widths, heights=heights, top=textGrob(title, gp=gpar(cex=1.3)),
                                      left=tgt.label, bottom=src.label))
  }
  #if (is.character(vertex.attr)) {
  # reorder the vertices so that vertices in same group are together

  #}
  #if (is.character(vertex.attr)) {
  #  vertices <- colnames(adj)
  #  jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))
  #  V.gr <- factor(get.vertex.attribute(g, vertex.attr))  # group the vertices by their attribute
  #  un.vertices <- levels(V.gr)
  #  attr.colors <- jet.colors(length(un.vertices))
  #  for (i in 1:length(attr.colors)) {
  #    vertex.num <- min(which())
  #    plot.adj <- plot.adj +
  #      geom_rect()
  #  }
  #  plot.adj <- plot.adj +
  #    geom_tile(data=adj.data, aes(x=Source, y=Target, fill=Fill1), show.legend=TRUE, alpha=0.2) +
  #    geom_tile(data=adj.data, aes(x=Source, y=Target, fill=Fill2), show.legend=FALSE, alpha=0.2)
  #}
  return(plot.adj)
}

gs.plot.sbm <- function(g, title="", edge.attr="dMRI", src.label="Source Community", tgt.label="Target Community") {
  V.comm <- V(g)$community
  names(V.comm) <- names(V(g))
  Vlist <- do.call(c, V.comm)  # get the full list of vertices ordered by community
  Vlist.comm <- do.call(c, lapply(names(V.comm), function(comm) rep(comm, length(V.comm[[comm]]))))
  P.mtx <- matrix(0, nrow=length(Vlist.comm), ncol=length(Vlist.comm))
  colnames(P.mtx) <- Vlist.comm
  rownames(P.mtx) <- Vlist.comm
  edge.vals <- get.edge.attribute(g, edge.attr)
  g.es <- E(g)
  for (i in 1:length(g.es)) {
    vtxs <- ends(g, g.es[i])
    P.mtx[rownames(P.mtx) == vtxs[1], colnames(P.mtx) == vtxs[2]] <- edge.vals[i]
  }
  rownames(P.mtx) <- Vlist; colnames(P.mtx) <- Vlist
  melted.p <- melt(P.mtx); names(melted.p) <- c("source", "target", edge.attr)
  vertices.data <- data.frame(id=Vlist, community=Vlist.comm)
  sbm.adjg <- graph_from_data_frame(melted.p, vertices=vertices.data)
  hmap.sbm <- gs.plot.heatmap(sbm.adjg, title=title, src.label=src.label, tgt.label=tgt.label,
                              degree=FALSE, edge.attr = edge.attr)
  return(hmap.sbm)
}


gs.plot.latent_pos <- function(X, title="",x.label="Dimension", y.label="Vertex",
                         font.size=NULL) {
  cvec <- c("#020202")
  plot.dat <- data.frame()
  adj.data <- melt(X)  # melt the graph to a data-frame with row and colnames preserved
  colnames(adj.data) <- c("Vertex", "Dimension", "Weight")
  thm = list(theme_void(),
               guides(fill=FALSE),
               theme(plot.margin=unit(rep(0,4), "lines")))
  plot.adj <- ggplot(adj.data, aes(x=Dimension, y=Vertex, alpha=Weight)) +
    geom_tile(fill=cvec) +
    xlab(x.label) +
    ylab(y.label) +
    ggtitle(title) +
    theme_bw() +
    theme(rect=element_blank(), panel.grid=element_blank())
  return(plot.adj)
}
```

```{r}
mouse.plot <- gs.plot.heatmap(mouse.duke, title="(C) Mouse dMRI Connectome", src.label="",
                              tgt.label="", edge.attr="weight", edge.xfm="log10")
fly.plot <- gs.plot.heatmap(fly.left, title="(B) Fly Mushroom Body Connectome", src.label="",
                            tgt.label="", degree=TRUE)
celegans.plot <- gs.plot.grid(celegans.male, title="(A) Celegans Connectome", src.label="",
                              tgt.label="", edge.attr=c("chem", "gap"), degree=TRUE)
human.plot <- gs.plot.heatmap_overlay(human.mri, title="(D) Human MRI Connectome",
                                      src.label="", tgt.label="",
                                      edge.attr=c("fMRI", "dMRI"),
                                      degree=TRUE)
```

```{r}
grid.arrange(celegans.plot, fly.plot, mouse.plot, human.plot, nrow=2)
```

# SBM Plots

```{r}
human.dmri.sbm <- gs.sbm.fit(human.mri, community.attribute = "lobe", edge.attr="dMRI")
# human.dmri.sbm <- delete_vertices(human.dmri.sbm, "non-labeled")
human.dmri.sbm.plot <- gs.plot.sbm(human.dmri.sbm, edge.attr="dMRI", title="(D.ii) Human dMRI")
human.fmri.sbm <- gs.sbm.fit(human.mri, community.attribute = "lobe", edge.attr="fMRI")
# human.fmri.sbm <- delete_vertices(human.fmri.sbm, "non-labeled")
human.fmri.sbm.plot <- gs.plot.sbm(human.fmri.sbm, edge.attr="fMRI", title="(D.i) Human fMRI")

mouse.sbm <- gs.sbm.fit(mouse.duke, community.attribute = "level1", edge.attr="weight")
mouse.sbm <- delete_vertices(mouse.sbm, "white matter")
mouse.sbm.plot <- gs.plot.sbm(mouse.sbm, edge.attr="weight", title="(C) Mouse dMRI")

fly.sbm <- gs.sbm.fit(fly.left, community.attribute = "type", edge.attr=NULL)
fly.sbm.plot <- gs.plot.sbm(fly.sbm, edge.attr="weight", title="(B) Fly Mushroom Body")

celegans.gap.sbm <- gs.sbm.fit(celegans.male, community.attribute = "type", edge.attr="gap")
celegans.chem.sbm <- gs.sbm.fit(celegans.male, community.attribute = "type", edge.attr="chem") 
celegans.gap.sbm.plot <- gs.plot.sbm(celegans.gap.sbm, edge.attr="gap", title="(A.ii) Celegans Gap")
celegans.chem.sbm.plot <- gs.plot.sbm(celegans.chem.sbm, edge.attr="chem", title="(A.i) Celegans Chemical")
```

```{r}
grid.arrange(celegans.chem.sbm.plot, celegans.gap.sbm.plot, fly.sbm.plot, mouse.sbm.plot, human.fmri.sbm.plot, human.dmri.sbm.plot, nrow=3)
```

# Spectral Embeddings

```{r}
human.dmri.dim <- gs.dim.select(human.mri, edge.attr="dMRI", plot=TRUE)
human.fmri.dim <- gs.dim.select(human.mri, edge.attr="fMRI", plot=TRUE)
human.dmri.ase <- gs.embed.ase(human.mri, edge.attr="dMRI")
human.fmri.ase <- gs.embed.ase(human.mri, edge.attr="fMRI")
mouse.dim <- gs.dim.select(mouse.duke, edge.attr="weight", plot=TRUE)
mouse.ase <- gs.embed.ase(mouse.duke, edge.attr="weight")
fly.dim <- gs.dim.select(fly.left, edge.attr=NULL, plot=TRUE)
fly.ase <- gs.embed.ase(fly.left)
celegans.chem.dim <- gs.dim.select(celegans.male, edge.attr="chem", plot=TRUE)
celegans.gap.dim <- gs.dim.select(celegans.male, edge.attr="gap", plot=TRUE)
celegans.chem.ase <- gs.embed.ase(celegans.male, edge.attr="chem")
celegans.gap.ase <- gs.embed.ase(celegans.male, edge.attr="gap")
```

```{r}
human.dmri.ase.plot <- gs.plot.latent_pos(human.dmri.ase$X)
human.dmri.ase.plot <- human.dmri.ase.plot + theme(legend.position=NaN)
human.dmri.ase.plot.comb <- arrangeGrob(human.dmri.dim$plot + ggtitle("(D.ii) Human dMRI") +
                                        theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")),
                                        human.dmri.ase.plot +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")), nrow=2, heights=c(0.3, 0.7))
human.fmri.ase.plot <- gs.plot.latent_pos(human.fmri.ase$X)
human.fmri.ase.plot <- human.fmri.ase.plot + theme(legend.position=NaN)
human.fmri.ase.plot.comb <- arrangeGrob(human.fmri.dim$plot + ggtitle("(D.i) Human fMRI") +
                                        theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")),
                                     human.fmri.ase.plot +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")), nrow=2, heights=c(0.3, 0.7))


mouse.ase.plot <- gs.plot.latent_pos(mouse.ase$X)
mouse.ase.plot <- mouse.ase.plot + theme(legend.position=NaN)
mouse.ase.plot.comb <- arrangeGrob(mouse.dim$plot + ggtitle("(C) Mouse dMRI") +
                                     theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()),
                                   mouse.ase.plot +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")), nrow=2, heights=c(0.3, 0.7))

fly.ase.plot <- gs.plot.latent_pos(fly.ase$X)
fly.ase.plot <- fly.ase.plot + theme(legend.position=NaN)
fly.ase.plot.comb <- arrangeGrob(fly.dim$plot + ylab("") + ggtitle("(B) Fly Mushroom body") +
                                     theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")),
                                   fly.ase.plot +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")), nrow=2, heights=c(0.3, 0.7))

celegans.chem.ase.plot <- gs.plot.latent_pos(celegans.chem.ase$X)
celegans.chem.ase.plot <- celegans.chem.ase.plot + theme(legend.position=NaN)
celegans.chem.ase.plot.comb <- arrangeGrob(celegans.chem.dim$plot + ggtitle("(A.i) Celegans Chemical") +
                                     theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()),
                                   celegans.chem.ase.plot, nrow=2, heights=c(0.3, 0.7))
celegans.gap.ase.plot <- gs.plot.latent_pos(celegans.gap.ase$X)
celegans.gap.ase.plot <- celegans.gap.ase.plot + theme(legend.position=NaN)
celegans.gap.ase.plot.comb <- arrangeGrob(celegans.gap.dim$plot + ggtitle("(A.ii) Celegans Gap") +
                                     theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")),
                                   celegans.gap.ase.plot +
                                     theme(axis.title.x=element_text(color="#FFFFFF"),
                                           axis.title.y=element_text(color="#FFFFFF")), nrow=2, heights=c(0.3, 0.7))
```

```{r}
grid.arrange(celegans.chem.ase.plot.comb, celegans.gap.ase.plot.comb, fly.ase.plot.comb, mouse.ase.plot.comb,
             human.fmri.ase.plot.comb, human.dmri.ase.plot.comb, nrow=3)
```

```{r}
rownames(fly.ase$X) <- V(fly.left)$type
gs.plot.matrix.pairs(fly.ase$X, pt.label="Type")

rownames(celegans.chem.ase$X) <- V(celegans.male)$type
gs.plot.matrix.pairs(celegans.chem.ase$X, pt.label="Type")

rownames(celegans.gap.ase$X) <- V(celegans.male)$type
gs.plot.matrix.pairs(celegans.gap.ase$X, pt.label="Type")

rownames(mouse.ase$X) <- V(mouse.duke)$level1
gs.plot.matrix.pairs(mouse.ase$X, pt.label="Region")

rownames(human.dmri.ase$X) <- V(human.mri)$type
gs.plot.matrix.pairs(human.dmri.ase$X, pt.label="Region")
```

Figure 3: Dimension1 vs Dimension2
color coded by attribute

Color vertices by their label
pairs plots
LLG paper
omni plots; similar to pairs plots but have a line joining matched vertices
fly left; fly right with matched vertices btwn the 2

separate mouse connectome by left and right hemisphere
use version of human lobe-wise assignment with desikan instead of brodmann
produce pairs plots for all 6
finish omni plots for fly, celegans, and human
